---
title: "Week 1 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv) and d the data into R. The code book, describing the variable names is [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf). How many properties are worth $1,000,000 or more?

```{r}
download.file(url = 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv',
              destfile = paste0(getwd(), '/quiz1_data1.csv'),
              method = 'curl', quiet = T)
raw <- read.csv('quiz1_data1.csv')
#VAL: property value
#24: 1000000+
table(raw$VAL == 24)
```

2. Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?
 - Each tidy data table contains information about only one type of observation
 - Tidy data has one observation per row.
 - **Tidy data has one variable per column.**
 - Each variable in a tidy data set has been transformed to be interpretable.

3. Download the Excel spreadsheet on Natural Gas Aquisition Program [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx). Read rows 18-23 and columns 7-15 into R and assign the result to a variable called `dat`. What is the value of `sum(dat$Zip*dat$Ext,na.rm=T)`
```{r}
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx', 
              destfile = paste0(getwd(), '/quiz1_data2.xlsx'), 
              method = 'curl', quiet = T)

library(openxlsx)
dat <- read.xlsx('quiz1_data2.csv', sheet = 1,rows = 18:23, cols = 7:15)
sum(dat$Zip*dat$Ext,na.rm=T)
```

4. Read the XML data on Baltimore restaurants from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml). How many restaurants have zipcode 21231?
```{r}
library(XML)
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml', 
              destfile = paste0(getwd(), '/quiz1_data3.xml'), 
              method = 'curl', quiet = T)
  #can't use xmlTreeParse directly b/c this is from 'https' (not supported by the package anymore!)
Q4data <- xmlTreeParse('quiz1_data3.xml', useInternalNodes = T)
Q4rootNode <- xmlRoot(Q4data)
  #wrapper element for the entire document
  #to gain access to the document
length(Q4rootNode[xpathSApply(Q4rootNode, '//zipcode', xmlValue) == 21231])
  #xpathSApply --> extract the values that have the tags 'zipcode'
  #find how many zipcodes are 21231
```

5. The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv). Using the `fread()` command load the data into an R object `DT`.

The following are ways to calculate the average value of the variable `pwgtp15` broken down by sex. Using the data.table package, which will deliver the fastest user time?
```{r}
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv', 
              destfile = paste0(getwd(), '/quiz1_data4.csv'), 
              method = 'curl', quiet = T)
library(data.table)
DT <- fread(input = 'quiz1_data4.csv', sep = ',')
DT[,mean(pwgtp15),by=SEX]
  #only this uses DT
```

