---
title: "Week 2 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Register an application with the Github API [here](https://github.com/settings/applications). Access the API to get information on your instructors repositories  
  (hint: this is the url you want <https://api.github.com/users/jtleek/repos>).  
  Use this data to find the time that the datasharing repo was created. What time was it created?  
  (Hint: [This tutorial](https://github.com/hadley/httr/blob/master/demo/oauth2-github.r) may be useful. You may also need to run the code in the base R package and not R studio.)

```{r}
#create a new app with the Github development team - to get the client Id and token
#https://github.com/settings/applications/670785
#Fill out application name; url - github homepage
clientId <- '1c038dd7df6a509afe89'
clientTkn <- 'bd222cfcf1b956faa4cb24b0ab3b1589d84294f6'
library(httr)
myapp <- oauth_app('github', key = clientId, secret = clientTkn)
  #start application authentication process
sig <- sign_oauth1.0(myapp, token = clientId, token_secret = clientTkn)
  #create a sign in credential
Q1json <- GET('https://api.github.com/users/jtleek/repos', sig)
  #get the page (json data) that corresponds to the url with your sign in credentials
json1 <- content(Q1json) 
  #extract the json file with the content
  #this is usually hard to read
library(jsonlite)
Q1data <- jsonlite::fromJSON(toJSON(json1)) 
  #creates a dataframe from the json data - this is easier to read
colnames(Q1data)
Q1data[1:2, 1:3]
  #read the first couple of row/columns of the dataframe
Q1data[Q1data$full_name == 'jtleek/datasharing', ]$created_at
```

 - 2012-06-20T18:39:06Z
 - 2014-03-05T16:11:46Z
 - 2014-02-06T16:13:11Z
 - **2013-11-07T13:25:07Z**

2. The sqldf package allows for execution of SQL commands on R data frames. We will use the sqldf package to practice the queries we might send with the dbSendQuery command in RMySQL. Download the [American Community Survey data](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv) and load it into an R object called `acs`. Which of the following commands will select only the data for the probability weights pwgtp1 with ages less than 50?
```{r, message=FALSE, warning=FALSE}
download.file(url = 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv',
              dest = paste0(getwd(), '/quiz2_data1.csv'),
              method = 'curl', quiet = T)
acs <- read.csv('quiz2_data1.csv')
library(sqldf)
head(sqldf('select pwgtp1 from acs where AGEP < 50'))
```
 - sqldf("select * from acs where AGEP < 50 and pwgtp1")
 - sqldf("select pwgtp1 from acs")
 - **sqldf("select pwgtp1 from acs where AGEP < 50")**
 - sqldf("select * from acs")

3. Using the same data frame you created in the previous problem, what is the equivalent function to unique(acs$AGEP)
```{r}
head(sqldf("select distinct AGEP from acs"))
```
 - **sqldf("select distinct AGEP from acs")**
 - sqldf("select distinct pwgtp1 from acs")
 - sqldf("select unique AGEP from acs")
 - sqldf("select unique AGEP from acs")


4. How many characters are in the 10th, 20th, 30th and 100th lines of HTML from [this page](http://biostat.jhsph.edu/~jleek/contact.html). (Hint: the `nchar()` function in R may be helpful)
```{r}
Q4data <- readLines('http://biostat.jhsph.edu/~jleek/contact.html')
c(nchar(Q4data[10]), nchar(Q4data[20]), nchar(Q4data[30]), nchar(Q4data[100]))
```
 - 43 99 7 25
 - **45 31 7 25**
 - 45 31 7 31
 - 45 92 7 2
 - 45 0 2 2
 - 45 31 2 25
 - 43 99 8 6

5. Read [this data](https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for) set into R and report the sum of the numbers in the fourth of the nine columns. (Hint: this is a fixed width file format)
```{r}
Q5data <- read.fwf(file=url("https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for"), 
                   widths = c(10, 9, 9, 4, 5, 8, 5, 8, 5), skip = 4)
sum(Q5data$V4)
```
 - 101.83
 - 222243.1
 - 35824.9
 - 28893.3
 - **32426.7**
 - 36.5