---
title: "Week 1 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Consider influenza epidemics for two parent heterosexual families. Suppose that the
probability is 17% that at least one of the parents has contracted the disease. The
probability that the father has contracted influenza is 12% while the probability that both
the mother and father have contracted the disease is 6%. What is the probability that the
mother has contracted influenza?

```
A = Mother; B = Father
P(A or B) = 17%
P(B) = 12%
P(A and B) = 6%
P(A or B) = P(A) + P(B) - P(A and B)
17% = P(A) + 12% - 6%
```

```{r}
17-12+6
```

  - **11%**
  - 6%
  - 5%
  - 17%

2. A random variable, `X` is uniform, a box from 0 to 1 of height 1. (So that its density is
`f(x) = 1` for `0 <= x <= 1`.) What is its 75th percentile?  
```{r}
qunif(0.75)
```
  - 0.25
  - 0.10
  - **0.75**
  - 0.50

3. You are playing a game with a friend where you flip a coin and if it comes up heads you
give her `X` dollars and if it comes up tails she gives you `Y` dollars. The probability that the
coin is heads is p (some number between `0` and `1`.) What has to be true about `X` and `Y` to
make so that both of your expected total earnings is `0`. The game would then be called
“fair”.

```
Expected earning: -pX + (1-p)Y = 0
pX = (1-p)Y
```
  - `X = Y`
  - **`p/(1-p) = Y/X`**
  - `p = X/Y`
  - `p/(1-p) = X/Y`
  
4. A density that looks like a normal density (but may or may not be exactly normal) is
exactly symmetric about zero. (Symmetric means if you flip it around zero it looks the
same.) What is its median?
  - The median must be 1.
  - **The median must be 0.**
  - We can't conclude anything about the median.
  - The median must be different from the mean
  5. Consider the following PMF shown below in R
```{r}
x <- 1:4
p <- x/sum(x)
temp <- rbind(x, p)
rownames(temp) <- c("X", "Prob")
temp
```

   What is the mean?
```{r}
sum(x * p)
```
   - 1
   - 2
   - **3**
  - 4

6. [A web site](www.medicine.ox.ac.uk/bandolier/band64/b64-7.html) for home pregnancy
tests cites the following: “When the subjects using the test were women who collected
and tested their own samples, the overall sensitivity was 75%. Specificity was also low, in
the range 52% to 75%.” Assume the lower value for the specificity. Suppose a subject has
a positive test and that 30% of women taking pregnancy tests are actually pregnant. What
number is closest to the probability of pregnancy given the positive test?

```
P(Preg|+) = (P(+|Preg) * P(Preg)) / (P(+|Preg) * P(Preg) + P(+|NotPreg) * P(NotPreg))
```
```{r}
(0.75*0.3) / (0.75*0.3 + (1-0.52)*(1-0.3))
```

  - 30%
  - 10%
  - **40%**
  - 20%
