---
title: "Week 2 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. What is the variance of the distribution of the average an IID draw of `n` observations from
a population with mean `μ` and variance `σ`?

    `Var(X) = σ^2/n`
  
  - **`σ^2/n`**
  - `σ/n`
  - `σ^2`
  - `2σ/sqrt(n)`

2. Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally
distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the
probability that a random 35-44 year old has a DBP less than 70?
```{r}
pnorm(70, mean = 80, sd = 10)
```

  - **16%**
  - 32%
  - 22%
  - 8%

3. Brain volume for adult women is normally distributed with a mean of about 1,100 cc for
women with a standard deviation of 75 cc. What brain volume represents the 95th
percentile?
```{r}
qnorm(0.95, mean = 1100, sd = 75)
```

  - approximately 1175
  - approximately 977
  - **approximately 1223**
  - approximately 1247

4. Refer to the previous question. Brain volume for adult women is about 1,100 cc for
women with a standard deviation of 75 cc. Consider the sample mean of 100 random
adult women from this population. What is the 95th percentile of the distribution of that
sample mean?
```{r}
qnorm(0.95, mean = 1100, sd = 75/sqrt(100))
```
  - approximately 1088 cc
  - **approximately 1112 cc**
  - approximately 1115 cc
  - approximately 1110 cc

5. You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?
`Choose(5, 4) * 2^-5 + Choose(5, 5) * 2^-5`
```{r}
pbinom(3, size = 5, prob = 0.5, lower.tail = FALSE)
```

  - 12%
  - **19%**
  - 3%
  - 6%

6. The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific
population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They
are not normally distributed. Give your best estimate of the probability that a sample
mean RDI of 100 people is between 14 and 16 events per hour?
```{r}
sem <- 10/sqrt(100)
pnorm(16, mean = 15, sd = sem) - pnorm(14, mean = 15, sd = sem)
```

  - 95%
  - 47.5%
  - 34%
  - **68%**

7. Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?
  - 0.75
  - **0.5**
  - 0.10
  - 0.25

8. The number of people showing up at a bus stop is assumed to be
Poisson with a mean of `5` people per hour. You watch the bus
stop for 3 hours. About what's the probability of viewing 10 or fewer people?
```{r}
ppois(10, lambda = 15)
```

  - **0.12**
  - 0.08
  - 0.06
  - 0.03

