---
title: "Week 1 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Which of the following are components in building a machine learning algorithm?

  - **Asking the right question.**
  - Statistical inference
  - Machine learning
  - Training and test sets
  - Artificial intelligence

2. Suppose we build a prediction algorithm on a data set and it is 100% accurate on that data set. Why might the algorithm not work well if we collect a new data set?

  - We have used neural networks which has notoriously bad performance.
  - We may be using bad variables that don't explain the outcome.v
  - **Our algorithm may be overfitting the training data, predicting both the signal and the noise.**
  - We have too few predictors to get good out of sample accuracy.

3. What are typical sizes for the training and test sets?

  - 20% training set, 80% test set.
  - 90% training set, 10% test set
  - 100% training set, 0% test set.
  - **60% in the training set, 40% in the testing set.**

4. What are some common error rates for predicting binary variables (i.e. variables with two possible values like yes/no, disease/normal, clicked/didn't click)? Check the correct answer(s).

  - Median absolute deviation
  - Root mean squared error
  - Correlation
  - **Accuracy**
  - R^2

5. Suppose that we have created a machine learning algorithm that predicts whether a link will be clicked with 99% sensitivity and 99% specificity. The rate the link is clicked is 1/1000 of visits to a website. If we predict the link will be clicked on a specific visit, what is the probability it will actually be clicked?

`P(Click|Visits) = (P(Visits|Click) * P(Click)) / (P(Visits|Click) * P(Click) + P(Visits|NoClicks) * P(NoVisits))`

```{r}
(0.99 * (1/1000)) /  (0.99 * (1/1000) + (1-0.99) * (1-1/1000))
```

  - 50% 
  - 99.9%
  - 99%
  - **9%**