---
title: "Week 1 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Which of the following are absolutely necessary for creating a functioning shiny app? (Check all that apply)
  - A server.R file that sets configuration options for hosting the App
  - **A server.R file containing a call to shinyServer()**
  - **A ui.R file containing a call to shinyUI()**
  - A shiny.R file containing calls to shinyServer() and shinyUI()
  - A ui.R file that contains information about the CSS and styling of the App

2. What is incorrect about the following syntax in ui.R?
```{r, eval = FALSE}
library(shiny)
shinyUI(pageWithSidebar(  
  headerPanel("Data science FTW!"),  
  sidebarPanel(    
     h2('Big text')    
     h3('Sidebar')  
  ),  
  mainPanel(      
       h3('Main Panel text')  
  )
))
```

  - Missing comma after the h3 command
  - The h3 command should be an h2 command
  - **Missing a comma in the sidebar panel**
  - The h2 command does not take text arguments

3. Consider the following in ui.R
```{r, eval = FALSE}
shinyUI(pageWithSidebar(  
   headerPanel("Example plot"),  
   sidebarPanel(    
     sliderInput('mu', 'Guess at the mu',value = 70, min = 60, max = 80, step = 0.05,)  ), 
   mainPanel(    
     plotOutput('newHist')  
   )
))
```

   And the following in server.R
```{r, eval = FALSE}
library(UsingR)
data(galton)

shinyServer(  
    function(input, output) {    
       output$myHist <- renderPlot({      
          hist(galton$child, xlab='child height', col='lightblue',main='Histogram')      
          mu <- input$mu      
          lines(c(mu, mu), c(0, 200),col="red",lwd=5)      
          mse <- mean((galton$child - mu)^2)      
          text(63, 150, paste("mu = ", mu))      
          text(63, 140, paste("MSE = ", round(mse, 2)))      
          })      }
)
```

   Why isn't it doing what we want? (Check all that apply.)

  - **The server.R output name isn't the same as the plotOutput command used in ui.R.**
  - It shoud be `mu <_ input$mean` in server.R
  - The limits of the slider are set incorrectly and giving an error.
  - The phrase "Guess at the mu value" should say "mean" instead of "mu"

4. What are the main differences between creating a Shiny Gadget and creating a regular Shiny App? (Check all that apply)

  - Shiny Gadgets can be run on a user’s personal computer, unlike a regular Shiny App which needs to be hosted online.
  - Shiny Gadgets are specially designed for use on mobile phones and tablet computers.
  - **Shiny Gadgets are designed to have small user interfaces that fit on one page.**
  - **Shiny Gadgets are designed to be used by R users in the middle of a data analysis.**
  - Shiny Gadgets are smaller programs and therefore run faster than Shiny Apps.

5. Consider the following R script:
```{r, eval = FALSE}
library(shiny)
library(miniUI)

pickXY <- function() {
  ui <- miniPage(
    gadgetTitleBar("Select Points by Dragging your Mouse"),
    miniContentPanel(
      plotOutput("plot", height = "100%", brush = "brush")
    )
  )

  server <- function(input, output, session) {
      output$plot <- renderPlot({
        plot(data_frame$X, data_frame$Y, main = "Plot of Y versus X",
           xlab = "X", ylab = "Y")
      })
      observeEvent(input$done, {
        stopApp(brushedPoints(data_frame, input$brush,
                          xvar = "X", yvar = "Y"))
      })
  }

  runGadget(ui, server)
}

my_data <- data.frame(X = rnorm(100), Y = rnorm(100))

pickXY(my_data)
```

   Why isn’t it doing what we want?

  - The wrong column names are passed to brushedPoints()
  - The input data is defined in such a way that it is not compatible with pickXY()
  - The call to plot() references the column names of the data frame in the wrong order.
  - **No arguments are defined for pickXY()**