---
title: "Week 1 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. R was developed by statisticians working at
 - Johns Hopkins University
 - Microsoft
 - Harvard University
 - **The University of Auckland**

2. The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition? Select all that
apply.
 - **The freedom to sell the software for any price.**
 - The freedom to redistribute copies so you can help your neighbor.
 - The freedom to run the program, for any purpose.
 - **The freedom to prevent users from using the software for undesirable purposes.**
 - The freedom to improve the program, and release your improvements to the public, so that the whole community benefits.
 - **The freedom to restrict access to the source code for the software.**
 - The freedom to study how the program works, and adapt it to your needs.

3. In R the following are all atomic data types EXCEPT: (Select all that apply)
 - **data frame**
 - **matrix**
 - complex
 - numeric
 - character
 - **table**
 - logical
 - **array**
 - integer
 - **list**

4. If I execute the expression `x <- 4` in R, what is the class of the object x as determined by the `class()` function?
```{r}
x <- 4
class(x)
```

 - matrix
 - real
 - complex
 - vector
 - **numeric**
 - integer
 - list

5. What is the class of the object defined by the expression `x <- c(4, "a", TRUE)`?
```{r}
x <- c(4, "a", TRUE)
class(x)
```

 - logical
 - integer
 - **character**
 - numeric
 - mixed

6. If I have two vectors `x <- c(1, 3, 5)` and `y <- c(3, 2, 10)`, what is produced by the expression `cbind(x, y)`?
```{r}
x <- c(1, 3, 5)
y <- c(3, 2, 10)
cbind(x, y)
```

 - a 2 by 2 matrix
 - a 3 by 3 matrix
 - **a matrix with 2 columns and 3 rows**
 - a 2 by 3 matrix
 - a vector of length 2
 - a vector of length 3

7. A key property of vectors in R is that
 - a vector cannot have attributes like dimensions
 - the length of a vecotr must be less than 32,768
 - elements of a vector can only be character or numeric
 - elements of a vector can be of different classes
 - **elements of a vector all must be of the same class**

8. Suppose I have a list defined as `x <- list(2, "a", "b", TRUE)`. What does `x[[2]]` give me? Select all that apply.
```{r}
x <- list(2, "a", "b", TRUE)
x[[2]]
```

 - a list containing the number 2 and the letter "a".
 - **a character vector of length 1.**
 - **a character vector containing the letter "a".**
 - a character vector with the elements "a" and "b".
 - a list containing character vector with the letter "a".

9. Suppose I have a vector `x <- 1:4` and a vector `y <- 2`. What is produced by the expression `x + y`?
```{r}
x <- 1:4
y <- 2
x + y
```

 - a numeric vector with elements 3, 2, 3, 6.
 - an integer vector with elements 3, 2, 3, 6.
 - an integer vector with elements 3, 2, 3, 4.
 - **a numeric vector with elements 3, 4, 5, 6.**

10. Suppose I have a vector `x <- c(3, 5, 1, 10, 12, 6)` and I want to set all elements of this vector that are less than 6 to be equal to zero. What R code achieves this? Select all that apply.
```{r}
x <- c(3, 5, 1, 10, 12, 6)
x[x %in% 1:5] <- 0
x
x[x < 6] <- 0
x
x[x <= 5] <- 0
x
```

 - `x[x > 6] <- 0`
 - `x[x == 0] <- 6`
 - `x[x == 6] <- 0`
 - `x[x >= 6] <- 0`
 - `x[x < 6] == 0`
 - **`x[x %in% 1:5] <- 0`**
 - **`x[x < 6] <- 0`**
 - **`x[x <= 5] <- 0`**
 - `x[x > 0] <- 6`
 - `x[x == 0] < 6`
 - `x[x != 6] <- 0`
 
11. Use the Week 1 Quiz Data Set to answer questions 11-20. In the dataset provided for this Quiz, what are the column names of the dataset?
```{r}
raw <- read.csv('hw1_data.csv')
colnames(raw)
```
 - **Ozone, Solar.R, Wind, Temp, Month, Day**
 - Month, Day, Temp, Wind
 - Ozone, Solar.R, Wind
 - 1, 2, 3, 4, 5, 6
 
 12. Extract the first 2 rows of the data frame and print them to the console. What does the output look like?
 
```{r}
raw[1:2, ]
```
 - **1 Ozone Solar.R Wind Temp Month Day  
   2 1 41 190 7.4 67 5 1  
   3 2 36 118 8.0 72 5 2**
 - 1 Ozone Solar.R Wind Temp Month Day  
   2 1 7 NA 6.9 74 5 11  
   3 2 35 274 10.3 82 7 17
 - 1 Ozone Solar.R Wind Temp Month Day  
   2 1 18 224 13.8 67 9 17  
   3 2 NA 258 9.7 81 7 22
 - 1 Ozone Solar.R Wind Temp Month Day  
   2 1 9 24 10.9 71 9 14  
   3 2 18 131 8.0 76 9 29

13. How many observations (i.e. rows) are in this data frame?
```{r}
nrow(raw)
```
 - 45
 - 129
 - **153**
 - 160
 
 14. Extract the last 2 rows of the data frame and print them to the console. What does the output look like?
 
```{r}
raw[152:153, ]
```

15. What is the value of Ozone in the 47th row?
```{r}
raw$Ozone[47]
```

 - 18
 - 63
 - **21**
 - 34

16. How many missing values are in the Ozone column of this data frame?
```{r}
table(is.na(raw$Ozone))
```

 - 43
 - **37**
 - 78
 - 9

17. What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.
```{r}
mean(raw$Ozone, na.rm = T)
```

 - 31.5
 - 18.0
 - 53.2
 - **42.1**

18. Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?
```{r}
mean(subset(raw, Ozone > 31 & Temp > 90)$Solar.R)
```

 - **212.8**
 - 185.9
 - 334.0
 - 205.0

19. What is the mean of "Temp" when "Month" is equal to 6?
```{r}
mean(raw$Temp[raw$Month == 6], na.rm = T)
```

 - 90.2
 - 85.6
 - **79.1**
 - 75.3

20. What was the maximum ozone value in the month of May (i.e. Month is equal to 5)?
```{r}
max(raw$Ozone[raw$Month == 5], na.rm = T)
```

 - 18
 - 97
 - 100
 - **115**
