---
title: "Week 2 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Suppose I define the following function in R
```{r}
cube <- function(x, n){
  x^3
}
```
   What is the result of running the following in R after defining this function?
```{r}
cube(3)
```
 - An error is returned becuase 'n' is not specified in the call to 'cube'
 - **The number 27 is returned**
 - The users is prompted to specify the value of 'n'.
 - A warning is given with no value returned.

2. The following code will produce a warning in R. Why?
```{r}
x <- 1:10
if(x > 5) {
  x <- 0
}
```
 - There are no elements in 'x' that are greater than 5
 - The syntax of this R expression is incorrect.
 - The expression uses curly braces.
 - You cannot set 'x' to be 0 because 'x' is a vector and 0 is a scalar.
 - **'x' is a vector of length 10 and 'if' can only test a single logical statement.**

3. Consider the following function
```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
```
   If I then run in R the following, what value is returned?
```{r}
z <- 10
f(3)
```
 - 16
 - **10**
 - 7
 - 4

4. Consider the following expression:
```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```
    What is the value of 'y' after evaluating this expression?
 - 3
 - NA
 - 5
 - **10**

5. Consider the following R function
```{r}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}
```
    Which symbol in the above function is a free variable?
 - **f**
 - z
 - d
 - L
 - g

6. What is an environment in R?
 - **a collection of symbol/value pairs**
 - a list whose elements are all functions
 - an R package that only contains data
 - a special type of function
 
7. The R language uses what type of scoping rule for resolving free variables?
 - global scoping
 - **lexical scoping**
 - dynamic scoping
 - compilation scoping

8. How are free variables in R functions resolved?
 - The values of free variables are searched for in the environment in which the function was called
 - The values of free variables are searched for in the global environment
 - The values of free variables are searched for in the working directory
 - **The values of free variables are searched for in the environment in which the function was defined**
 
9. What is one of the consequences of the scoping rules used in R?
 - Functions cannot be nested
 - All objects can be stored on the disk
 - **All objects must be stored in memory**
 - R objects cannot be larger than 100 MB
 
10. In R, what is the parent frame?
 - It is the environment in which a function was defined
 - **It is the environment in which a function was called**
 - It is always the global environment
 - It is the package search list
