---
title: "Week 3 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Take a look at the 'iris' dataset that comes with R. The data can be loaded with the code:
```{r}
library(datasets)
data(iris)
```
A description of the dataset can be found by running
```{r}
?iris
```
There will be an object called 'iris' in your workspace. In this dataset, what is the mean of 'Sepal.Length' for the species virginica? **Please round your answer to the nearest whole number**. (Only enter the numeric result and nothing else.)
```{r}
round(mean(subset(iris, Species == 'virginica')$Sepal.Length, na.rm = T), 0) #using subset
```

2. Continuing with the 'iris' dataset from the previous Question, what R code returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', 'Petal.Length', and 'Petal.Width'?
```{r}
apply(iris[, 1:4], 2, mean)
```

 - `apply(iris[, 1:4], 1, mean)`
 - `rowMeans(iris[, 1:4])`
 - `colMeans(iris)`
 - `apply(iris, 2, mean)`
 - `apply(iris, 1, mean)`
 - **`apply(iris[, 1:4], 2, mean)`**

3. Load the 'mtcars' dataset in R with the following code
```{r}
library(datasets)
data(mtcars)
```
There will be an object names 'mtcars' in your workspace. You can find some information about the dataset by running
```{r}
?mtcars
```
How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)? Select all that apply.
```{r}
with(mtcars, tapply(mpg, cyl, mean))
tapply(mtcars$mpg, mtcars$cyl, mean)
sapply(split(mtcars$mpg, mtcars$cyl), mean)
```


 - `sapply(mtcars, cyl, mean)`
 - **`with(mtcars, tapply(mpg, cyl, mean))`**
 - `apply(mtcars, 2, mean)`
 - **`tapply(mtcars$mpg, mtcars$cyl, mean)`**
 - `lapply(mtcars, mean)`
 - **`sapply(split(mtcars$mpg, mtcars$cyl), mean)`**
 - `mean(mtcars$mpg, mtcars$cyl)`
 - `tapply(mtcars$cyl, mtcars$mpg, mean)`
 - `split(mtcars, mtcars$cyl)`

4. Continuing with the 'mtcars' dataset from the previous Question, what is the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars? (**Please round your final answer to the nearest whole number**. Only enter the numeric result and nothing else.)
```{r}
round(abs(mean(mtcars[mtcars$cyl == 8, ]$hp) - mean(mtcars[mtcars$cyl == 4, ]$hp)), 0)
round(abs(tapply(mtcars$hp, mtcars$cyl, mean)[1] - tapply(mtcars$hp, mtcars$cyl, mean)[3]), 0)
round(abs(sapply(split(mtcars$hp, mtcars$cyl)[1], mean) - sapply(split(mtcars$hp, mtcars$cyl)[3], mean)), 0)
```

5. If you run
```{r}
debug(ls)
```
what happens when you next call the 'ls' function?

 - **Execution of 'ls' will suspend at the beginning of the function and you will be in the browser.**
 - You will be prompted to specify at which line of the function you would like to suspend execution and enter the browser.
 - Execution of the 'ls' function will suspend at the 4th line of the function and you will be in the browser.
 - The 'ls' function will return an error.
