---
title: "Week 1 Assignment: Plotting Electric Power Consumption"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
This assignment uses data from the 
[UC Irvine Machine Learning Repository](http://archive.ics.uci.edu/ml/), a popular repository for machine learning datasets. In particular, we will be using the "Individual household electric power consumption Data Set" which I have made available on the course web site:

  - **Dataset**: [Electric power consumption](https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip) [20Mb]
  - **Description**:Measurements of electric power consumption in
one household with a one-minute sampling rate over a period of almost
4 years. Different electrical quantities and some sub-metering values
are available.

The following descriptions of the 9 variables in the dataset are taken
from the [UCI web site](https://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption)
  1. **Date**: Date in format dd/mm/yyyy
  2. **Time**: time in format hh:mm:ss
  3. **Global_active_power**: household global minute-averaged active power (in kilowatt)
  4. **Global_reactive_power**: household global minute-averaged reactive power (in kilowatt)
  5. **Voltage**: minute-averaged voltage (in volt)
  6. **Global_intensity**: household global minute-averaged current intensity (in ampere)
  7. **Sub_metering_1**: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
  8. **Sub_metering_2**: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
  9. **Sub_metering_3**: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.

## Loading the data
When loading the dataset into R, please consider the following:

  - The dataset has 2,075,259 rows and 9 columns. First
calculate a rough estimate of how much memory the dataset will require
in memory before reading into R. Make sure your computer has enough
memory (most modern computers should be fine).
  - We will only be using data from the dates 2007-02-01 and
2007-02-02. One alternative is to read the data from just those dates
rather than reading in the entire dataset and subsetting to those
dates.
  - You may find it useful to convert the Date and Time variables to
Date/Time classes in R using the `strptime()` and `as.Date()`
functions.
  - Note that in this dataset missing values are coded as `?`.

```{r, eval = FALSE}
download.file('https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip',
              destfile = './hw1_data.zip', method = 'curl', quiet = T)
unzip(zipfile = 'hw1_data.zip')
```

```{r}
rawData <- read.table('household_power_consumption.txt', header = T, stringsAsFactors = F,
                      na.strings = "?", sep = ';', quote = "", skip = 66636, nrows = 2880)
tail(rawData, 10)
colnames(rawData) <- c('Date', 'Time', 'GlobalActivePower', 'GlobalReactivePower', 'Voltage',
                       'GlobalIntensity', 'SubMetering1', 'SubMetering2', 'SubMetering3')
rawData$DateTime <- strptime(paste(rawData$Date, rawData$Time), format = '%d/%m/%Y %H:%M:%S')
```

## Making Plots

Our overall goal here is simply to examine how household energy usage
varies over a 2-day period in February, 2007. Your task is to
reconstruct the plots that can be found [here](https://github.com/rdpeng/ExData_Plotting1), all of which were constructed using the base plotting system.

For each plot you should

  - Construct the plot and save it to a PNG file with a width of 480
pixels and a height of 480 pixels.
  - Name each of the plot files as `plot1.png`, `plot2.png`, etc.
  - Create a separate R code file (`plot1.R`, `plot2.R`, etc.) that
constructs the corresponding plot, i.e. code in `plot1.R` constructs
the `plot1.png` plot. Your code file **should include code for reading
the data** so that the plot can be fully reproduced. You should also
include the code that creates the PNG file.
  - Add the PNG file and R code file to your git repository

### Plot 1
```{r}
#png('plot1.png', width = 480, height = 480)
hist(rawData$GlobalActivePower, col = 'red',
     main = 'Global Active Power',
     xlab = 'Global Active Power (kilowatts)')
#dev.off()
```

### Plot 2
```{r}
#png('plot2.png', width = 480, height = 480)
plot(x = rawData$DateTime, y = rawData$GlobalActivePower, 
     type = 'l', xlab = NA, ylab = 'Global Active Power (kilowatts)')
#dev.off()
```

### Plot 3
```{r}
#png('plot3.png', width = 480, height = 480)
plot(x = rawData$DateTime, y = rawData$SubMetering1, type = 'l',
     xlab = NA, ylab = 'Energy sub metering')
lines(x = rawData$DateTime, y = rawData$SubMetering2, col = 'red')
lines(x = rawData$DateTime, y = rawData$SubMetering3, col = 'blue')
legend('topright', 
       legend = c('Sub_metering_1', 'Sub_metering_2', 'Sub_metering_3'),
       col = c('black', 'red', 'blue'),
       lwd = 1)
#dev.off()
```

### Plot 4
```{r}
#png('plot4.png', width = 480, height = 480)
par(mfrow = c(2, 2)) 
  #plots from left to right, top to bottom (since you used mfrow vs mfcol)

#plot 4a = plot 1
plot(x = rawData$DateTime, y = rawData$GlobalActivePower, 
     type = 'l', xlab = NA, ylab = 'Global Active Power')

#plot 4b
plot(x = rawData$DateTime, y = rawData$Voltage, 
     type = 'l', xlab = 'datetime', ylab = 'Voltage')

#plot 4c = plot 3
plot(x = rawData$DateTime, y = rawData$SubMetering1, type = 'l',
     xlab = NA, ylab = 'Energy sub metering')
lines(x = rawData$DateTime, y = rawData$SubMetering2, col = 'red')
lines(x = rawData$DateTime, y = rawData$SubMetering3, col = 'blue')
legend('topright', 
       legend = c('Sub_metering_1', 'Sub_metering_2', 'Sub_metering_3'),
       col = c('black', 'red', 'blue'),
       lwd = 1, bty = 'n')

#plot 4d
plot(x = rawData$DateTime, y = rawData$GlobalReactivePower, type = 'l',
     xlab = 'datetime', ylab = 'Global_reactive_power')

#dev.off()
```
