---
title: "Week 3 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r}
suppressMessages(library(data.table))
nGramAll <- fread('predictionTableFull.csv')
source('nextWords.R')
```

1. Q1
```{r}
nextWords('The guy in front of me just bought a pound of bacon, a bouquet, and a case of', 5)
```
  - soda
  - pretzels
  - **beer**
  - cheese

2. Q2
```{r}
nextWords("You're the reason why I smile everyday. Can you follow me please? It would mean the", 5)
```
  - **world**
  - universe
  - best
  - most

3. Q3
```{r}
nextWords("Hey sunshine, can you follow me and make me the", 5)
```
  - saddest
  - smelliest
  - **happiest**
  - bluest

4. Q4
```{r}
result <- nextWords("Very early observations on the Bills game: Offense still struggling but the", 1000)
match(c('defense', 'crowd', 'players', 'referees'), result)
```
  - **defense**
  - *crowd*
  - players
  - referees
  
  - NOTE: incorrect prediction

5. Q5
```{r}
result <- nextWords("Go on a romantic date at the", 1000)
match(c('movies', 'grocery', 'beach', 'mall'), result)
```
  - movies
  - grocery
  - **beach**
  - mall

6. Q6
```{r}
nextWords("Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my", 5)
```
  - motocycle
  - phone
  - horse
  - **way**

7. Q7
```{r}
nextWords("Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some", 5)
```
  - years
  - weeks
  - **time**
  - thing

8. Q8
```{r}
nextWords("After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little", 5)
```

  - Try with only last two words
```{r}
result <- nextWords("his little", 15)
match(c('eyes', 'fingers', 'ears', 'toes'), result)
match(c('eye', 'finger', 'ear', 'toe'), result)
```

  - eyes
  - **fingers**
  - ears
  - toes
  
  - NOTE: incorrect prediction since my model uses the last three words to predicts

9. Q9
```{r}
result <- nextWords("Be grateful for the good times and keep the faith during the", 1000)
match(c('hard', 'worse', 'bad', 'sad'), result)
```
  - hard
  - worse
  - **bad**
  - sad
  
  - NOTE: unable to predict

10. Q10
```{r}
result <- nextWords("If this isn't the cutest thing you've ever seen, then you must be", 100)
match(c('insensitive', 'insane', 'callous', 'asleep'), result)
```
  - insensitive
  - **insane**
  - callous
  - asleep
  
  - NOTE: unable to predict
