---
title: "Week 4 Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r}
suppressMessages(library(data.table))
nGramAll <- fread('predictionTableFull.csv')
source('nextWords.R')
```

1. Q1
```{r}
result <- nextWords("When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd", 100)
match(c('sleep', 'give', 'die', 'eat'), result)
```
  - sleep
  - *give*
  - **die**
  - eat
  
  - NOTE: incorrect prediction

2. Q2
```{r}
result <- nextWords("Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his", 100)
match(c('marital', 'financial', 'spiritual', 'horticultural'), result)
```
  - **marital**
  - financial
  - spritual
  - horticultural
  
  - NOTE: unable to predict

3. Q3
```{r}
result <- nextWords("I'd give anything to see arctic monkeys this", 100)
match(c('morning', 'weekend', 'decade', 'month'), result)
```
  - *morning*
  - **weekend**
  - decade
  - month
  
  - NOTE: incorrect prediction

4. Q4
```{r}
result <- nextWords("Talking to your mom has the same effect as a hug and helps reduce your", 100)
match(c('happiness', 'sleepiness', 'hunger', 'stress'), result)
```
  - happiness
  - sleepiness
  - hunger
  - **stress**
  
5. Q5
```{r}
result <- nextWords("When you were in Holland you were like 1 inch away from me but you hadn't time to take a", 1000)
match(c('minute', 'look', 'walk', 'picture'), result)
```
  - minute
  - *look*
  - walk
  - **picture**
  
  - NOTE: incorrect prediction

6. Q6
```{r}
result <- nextWords("I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the", 1000)
match(c('account', 'case', 'incident', 'matter'), result)
```
  - account
  - *case*
  - incident
  - **matter**
  
  - NOTE: incorrect prediction (even though I think the answer is wrong)

7. Q7
```{r}
result <- nextWords("I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each", 1000)
match(c('finger', 'arm', 'toe', 'hand'), result)
```
  - finger
  - arm
  - toe
  - **hand**

8. Q8
```{r}
nextWords("Every inch of you is perfect from the bottom to the", 5)
```

  - middle
  - center
  - side
  - **top**

9. Q9
```{r}
result <- nextWords("I’m thankful my childhood was filled with imagination and bruises from playing", 1000)
match(c('outside', 'inside', 'weekly', 'daily'), result)
```
  - **outside**
  - insdie
  - weekly
  - daily

  - NOTE: unable to predict

10. Q10
```{r}
result <- nextWords("I like how the same people are in almost all of Adam Sandler's", 100)
match(c('movies', 'novels', 'stories', 'pictures'), result)
```
  - **movies**
  - novels
  - stories
  - pictures
  
  - NOTE: unable to predict